### 2015-09-30 committee_report/committee_report.tex
finished results and discussion

\subsection*{Abstract (50-words)}

Recent studies in breast cancer have shown that there are genetic signatures that are specific to obesity, and these signatures were further investigated in other cancer types (Creighton and fuentes mattei paper).
The obesity genes identified by Creighton \textit{et al.} did not correlate with any of the cancer types investigated.

@@

@@

Since the project was based on the results presented by Creighton \textit{et al.} (2012), that there were obesity-specific genetic signatures in breast cancer, the methods used in this project was applied to the raw data from their paper and the results were compared.

must not exceed 6 pages (with figures and references).

>>>>

\subsection*{50 Words Summary}

Recent studies in breast cancer have shown that there are obesity-specific genetic signatures, and these signatures were further investigated in other cancer types (Creighton and fuentes mattei paper).
The obesity genes identified by Creighton \textit{et al.} did not correlate with any of the cancer types investigated.

@@

%\newpage

@@

Since the project was based on the results presented by Creighton \textit{et al.} (2012), the methods used in this project was applied to the raw data from their paper and the results were compared to validate whether the methodolgy to be used in this project is applicable to other data.

After the Creighton \textit{et al.} data were normalised using RMA method, singular value decomposition was applied to the data to obtain the obesity metagene and plotted on a heatmap (fig1a).
The resulting heatmap showed that the overall expression of the obesity-specific genes correlated with the metagene.
To assess whether the metagene was actually correlating with the sample BMI, the metagene was plotted against the sample BMI status and sample BMI value (fig1b and c).
These plots clearly showed that the metagene was able to identify and separate the obese samples from the normal and overweight samples.
Together, these plots confirmed that the methodology applied to the data set was able to recapitulate the results presented by Creighton \textit{et al.}.

The metagenes for other cancer types were generated by using the obesity-specific genes from breast cancer.
Transformation matrix was applied to each cancer type to generate the metagene and was plotted on the heatmap (fig2).
All of the cancer types showed significant correlation with the metagene created from the transformation, where the high metagene values corresponded with high overall gene expression, and vice versa.

To confirm whether the association was in fact due to BMI of the samples, plots of metagene against BMI status and BMI value were created.
In contrast to the Creighton \textit{et al.} plots, no cancer types showed any significant separation between the obese samples and the normal/overweight samples.
These results suggest that the obesity-specific genes from breast cancer were not specific enough to be used as a genetic marker for other cancer types.

Although these current results do not show any significant correlation of the obesity-specific genes and the sample BMI in different cancers, there are many places for improvements.
The data processing of the raw RNA-seq data was carried out using a program written by myself, which could have processed the data incorrectly.
There are other programs to process these raw data, so the data processing procedure should be repeated and compared with the current results.

The obesity-specific genes in each cancer type was not checked with the BMI values of the samples.
The results presented here only showed the association of the metagene with the BMI value and status, and it would be good to double-check whether the expression of the genes show any correlation with the BMI of the samples.

For future experiments, differential expression analysis for each cancer type should be carried out to identify the genes that are dysregulated in obese samples but not in the normal or overweight data.
These genes can then be compared with the genes from other cancer types to identify genes that are  commonly up or down regulated in different cancers.
More recent paper by Fuentes-Mattei \textit{et al.} (2014) have also found obesity-specific genes in breast cancer, and these genes should be investigated as well.
